---
title: "Hands-on Exercise 4.2"
author: "Ha Duc Tien"
date: "May 02, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# 1. Getting Started

In this hands-on exercise, we will gain hands-on experience on using:

-   ggstatsplot package to create visual graphics with rich statistical information,

-   performance package to visualise model diagnostics, and

-   parameters package to visualise model parameters

[ggstatsplot](https://indrajeetpatil.github.io/ggstatsplot/) is an extension of ggplot2 package for creating graphics with details from statistical tests included in the information-rich plots themselves.

## 1.1 Install and launching R packages

The Code chunk below uses p_load() function of pacman package to check if tidyverse and ggstatsplot packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(ggstatsplot, tidyverse) 
```

## 1.2 Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
head(exam_data)
```


