---
title: "Hands-on Exercise 1 A Layered Grammar of Graphics: ggplot2 methods"
author: "Ha Duc Tien"
date: "April 12, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# 1. Getting Started

## 1.1 Install and launching R packages

The Code chunk below uses p_load() function of pacman package to check if tidyverse packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(tidyverse)
```

## 1.1 Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# 2. R Graphics VS ggplot

::: panel-tabset
## 2.1 R Graphics

```{r}
hist(exam_data$MATHS)
```

## 2.2 ggplot2

```{r}
ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=10, 
                 boundary = 100,
                 color="black", 
                 fill="grey") +
  ggtitle("Distribution of Maths scores")
```
:::

# 3. Plotting a simple bar chart

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) +
  geom_bar()
```

# 4. Plotting a simple dot plot

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) + 
  geom_dotplot(binwidth=2.5,
               dotsize = 0.5) + 
  scale_y_continuous(NULL,
                     breaks = NULL)
                     
```

# 5. Plotting a simple histogram plot

```{r}
ggplot(data=exam_data,
       aes(x = MATHS)) + 
  geom_histogram()       
```

## 5.1 Modifying a geometric object by changing aes()

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           fill = GENDER)) +
  geom_histogram(bins=20, 
                 color="grey30")
```

## 5.2 Modifying a geometric object by changing geom()

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20,            
                 color="black",      
                 fill="light blue")  
```

## 5.3 Using geom-density() to plots kernel density estimate

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS, 
           colour = GENDER)) +
  geom_density()
```

# 6. Plotting a simple box plot

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS,       
           x= GENDER)) +    
  geom_boxplot()            
```

## 6.1 Plotting a simple box plot using notched plot

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot(notch=TRUE)
```

# 7. Plotting a simple violin plot

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_violin()
```

# 8. Plotting a simple scatterplot

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point()            
```

## 8.1 Combining box plot and scatterplot

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot() +                    
  geom_point(position="jitter", 
             linewidth = 0.5)        
```

# 9. Essential Grammatical Elements in ggplot2: stat

## 9.1 Working with stat - the stat_summary() method

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  stat_summary(geom = "point",       
               fun="mean",         
               colour ="red",        
               size=4)               
```

## 9.2 Working with stat - the geom() method

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  geom_point(stat="summary",        
             fun ="mean",           
             colour ="red",          
             size=4)          
```

## 9.3 Adding a best fit curve on a scatterplot

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(linewidth=0.5)
```

### 9.3.1 Default smoothing method

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5)
```

# 10. Essential Grammatical Elements in ggplot2: Facets

## 10.1 Working with facet_wrap()

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_wrap(~ CLASS)
```

## 10.2 facet_grid() function

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_grid(~ CLASS)
```

# 11. Essential Grammatical Elements in ggplot2: Coordinates

## 11.1 Working with Coordinate

### 11.1.1 Flips the horizontal bar chart into vertical bar chart by using coord_flip()

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip()
```

## 11.2 Changing the y- and x-axis range

### 11.2.1 Fixed both the y-axis and x-axis range from 0-100.

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
```

# 12. Essential Grammatical Elements in ggplot2: themes

## 12.1 Working with theme

### 12.1.1 Plot a horizontal bar chart using theme_gray()

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_gray()
```

### 12.1.2 Plotted using theme_classic()

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_classic()
```

### 12.1.3 Plotted using theme_minimal()

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_minimal()
```

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_bw()
```
