---
title: "Hands-on Exercise 1"
author: "Ha Duc Tien"
---

# Getting Started

## Install and launching R packages

The Code chunk below uses p_load() function of pacman package to check if tidyverse packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(tidyverse)
```

## Importing the data

```{r}
exam_data <- read.csv("data/Exam_data.csv")
```

# Plotting a simple bar chart

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) +
  geom_bar()
```

# Plotting a simple dot plot
```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) + 
  geom_dotplot(binwidth=2.5,
               dotsize = 0.5) + 
  scale_y_continuous(NULL,
                     breaks = NULL)
                     
```

# Plotting a simple histogram plot

```{r}
ggplot(data=exam_data,
       aes(x = MATHS)) + 
  geom_histogram()       
```

## Modifying a geometric object by changing aes()

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           fill = GENDER)) +
  geom_histogram(bins=20, 
                 color="grey30")
```